\subsection{Threading Issues}

\subsubsection{Semantics of $fork$ and $exec$}

Should $fork$ duplicate all threads in the process,
or only the thread which called it? According to man, it duplicates the
process, with only one thread, that thread being the one that called it.

$exec$ replaces the entire running process, which includes all threads.


\subsubsection{Signal Handling}

Signals notify processes of events. A signal handler is used to process
incoming signals. This handler can either be the supplied default handler,
or a user-defined handler, which overrides the default handler.

Where should a signal be delivered in a multi-threaded program?

\begin{itemize}
    \item To the thread to which it applies.
    \item To every thread in the process.
    \item To certain threads.
    \item To a single thread which is assigned specifically to receive all signals.
\end{itemize}


\subsubsection{Thread Cancellation}

Thread cancellation allows a thread to be cancelled before it is finished.
This is generally done in one of two ways:
\begin{itemize}
    \item \textit{Asynchronous} --- Immediately terminates target thread.
    \item \textit{Deferred} --- Target thread periodically checks if it should be cancelled.
\end{itemize}

Invoking a cancellation \textit{requests} that a thread be cancelled. This request will
remain pending if the thread is not in a state in which it can be cancelled, e.g.\ the thread
may have disabled cancellations for a period, the request will remanin pending until it re-enables
cancellation.

The usual choice is Deferred. In POSIX \textit{pthreads}, cancellation will only occur
at a cancellation point, i.e.\ when $pthread_testcancel()$ is called. When a thread is cancelled
a cleanup handler is then invoked.


\subsubsection{Creating threads in Linux}

Threads are created in Linux using $clone$, with one of the four flags:
\begin{itemize}
    \item \textit{CLONE\_FS} --- Shares file-system information with the new thread.
    \item \textit{CLONE\_VM} --- The same memory space is shared.
    \item \textit{CLONE\_SIGHAND} --- Shares signal handlers with the new thread.
    \item \textit{CLONE\_FILES} --- The set of open files is shared with the new thread.
\end{itemize}