\section{What is interaction design?}

Interaction design includes a number of disciplines, including:
\begin{itemize}
    \item Computer science
    \item Psychology
    \item Sociology
    \item Engineering
    \item Product and Graphical design
    \item Ergonomics and Human Factors
    \item Cognitive science
\end{itemize}

\subsection{How to design}
When designing interactive systems, we should consider:
\begin{itemize}
    \item Who the users article
    \item What activities are being carried out
    \item Where the interaction is taking place
\end{itemize}

The interactions a user has with a system shoud be optimised in such a way that they match the user's activities, needs, and wants.
You must take into account what a user might be good / bad at, what might help people, what might provide
quality user experiences, things that people want, and user centered interaction.

\subsection{Usability goals}
\begin{itemize}
    \item Effective to use
    \item Efficient to use
    \item Safe to use 
    \item Has good utility 
    \item Easy to learn 
    \item Easy to remember
\end{itemize}

\subsection{User experience goals}
\begin{itemize}
    \item Satisfying
    \item Enjoyable
    \item Engaging
    \item Exciting
    \item Entertaining
    \item Rewarding
    \item Fulfilling
    \item Memorable
    \item Challenging
    \item Relaxing
    \item Comforting
    \item Fun
\end{itemize}

Based on a user's feelings when interacting with a UI.\ 

\subsection{Donald Norman's design principles}
\subsubsection{Visibility}
The user should not be expected to actively look for things, essential parts of the UI should be easily visible.

\subsubsection{Feedback}
Let the user know the status of the system, that an action has been performed successfully, or not.

\subsubsection{Constraints}
Restricts the user interactions that can take place, either permanently, or just for a given point in time.

\subsubsection{Mapping}
Make sure the relationship between a control and its effect is clear.

\subsubsection{Consistency}
Similar elements should have similar functions, thus allowing people to discover patterns they can
apply in new situations.

\subsubsection{Affordance}
Visual attributes should give clues to how a user should interact with something.